---
title: "p8105_hw3_yh3094"
author: "Yun He"
date: "October 5, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

# Problem 1

## Import the BRFSS data

```{r brfss_data_import}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
data("brfss_smart2010")
```

## Clean the BRFSS data

```{r brfss_data_clean}
brfss_data = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>% 
  mutate(response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor")))
```

## Pick out the states observed at 7 locations in 2002

```{r state_7_location}
brfss_data %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  summarize(n_location = n_distinct(locationdesc)) %>%
  filter(n_location == 7)
```

In 2002, CT, FL and NC were observed at 7 locations.

## Make a “spaghetti plot” 

The code chunk below makes a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.

```{r spaghetti_plot}
brfss_data %>% 
  group_by(year, locationabbr) %>% 
  summarize(n_location = n_distinct(locationdesc)) %>% 
  ggplot(aes(x = year, y = n_location, color = locationabbr)) +
  geom_line() +
  labs(title = "The number of locations in each state from 2002 to 2010")
```

From 2002 to 2010, the number of locations in most states was fewer than 10. In FL, the number of locations had the biggest change over time. Before 2007, the number of locations in FL was fewer than 10 each year, while in 2007, the number of locations in FL was up to 44. Then, the number of locations in FL decreased to fewer than 10 again. However, in 2010, the number of locations in FL increased to 41. 

## Make a table

The code chunk below makes a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

```{r excellent_table}
brfss_data %>% 
  filter(
    locationabbr == "NY", year %in% c("2002", "2006", "2010"), response == "Excellent"
  ) %>%
  group_by(year) %>% 
  summarise(
    mean_excellent = mean(data_value, na.rm = TRUE), 
    sd_excellent = sd(data_value, na.rm = TRUE)) %>%
  knitr::kable(digits = 1)
```

The mean of the proportion of “Excellent” responses across locations in NY State decreased in 2006 and then increased a little bit in 2010. The standard deviation of the proportion of “Excellent” responses across locations in NY State showed a decreasing trend from 2002 to 2010.

## Compute the average proportion in each response category for each year and state

```{r response_average}
brfss_data %>% 
  group_by(year, locationabbr, response) %>% 
  summarize(mean_response_proportion = round(mean(data_value, na.rm = TRUE), digits = 1)) 
```

## Make a five-panel plot

The code chunk below makes a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

```{r five_panel_plot}
brfss_data %>% 
  group_by(year, locationabbr, response) %>% 
  summarize(mean_response_proportion = mean(data_value, na.rm = TRUE)) %>% 
  ggplot(aes(x = factor(year), y = mean_response_proportion)) +
  facet_grid(~response) +
  geom_violin(aes(fill = factor(year))) +
  stat_summary(fun.y = median, geom = "point") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(x = "Year", y = "State-level averages", 
       title = "The distribution of the state-level response averages over time")
```

The state-level average proportion of “Very good” responses remained highest compared with that of other response categories over time. And the state-level average proportion of “Poor” responses remained lowest over time. For all response categories, these state-level averages didn't change much over time.

# Problem 2

## Import the Instacart data

```{r instacart_data_import}
data("instacart")
instacart = janitor::clean_names(instacart)
skimr::skim(instacart)
```

Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.

The Instacart data has `r nrow(instacart)` observations and `r ncol(instacart)` variables. Variables for `aisle`, `department`, `eval_set` and `product_name` are class `character` and other variables are class `integer`. The dataset contains products of 39,123 different names bought from 134 different aisles within 21 departments. `r round(sum(instacart$reordered == 1)/nrow(instacart), digits = 2)` of all items ordered have been ordered by the user in the past. Among those items which have been ordered before, `r (instacart %>% filter(reordered == 1) %>% count(product_name) %>% top_n(1))[[1]]` is the most popular. Most orders were placed at $14\pm4.2$ hours of the day. 

The first observation is `r instacart$product_name[1]` (product id: `r instacart$product_id[1]`) bought from the `r instacart$aisle[1]` aisle (aisle id: `r instacart$aisle_id[1]`) of the `r instacart$department[1]` department (department id: `r instacart$department_id[1]`) by the user whose id is `r instacart$user_id[1]`. The order (order id: `r instacart$order_id[1]`) was placed at `r instacart$order_hour_of_day[1]` on Thursday. The order sequence number for this user is `r instacart$order_number[1]`. This product was the first product added to cart by this user and has been ordered  before. `r instacart$days_since_prior_order[1]` days have past since the last order placed by this user. The evaluation set this order belongs in is `r instacart$eval_set[1]`. 

The ninth observation is `r instacart$product_name[9]` (product id: `r instacart$product_id[9]`) bought from the `r instacart$aisle[9]` aisle (aisle id: `r instacart$aisle_id[9]`) of the `r instacart$department[9]` department (department id: `r instacart$department_id[9]`) by the user whose id is `r instacart$user_id[9]`. The order (order id: `r instacart$order_id[9]`) was placed at `r instacart$order_hour_of_day[9]` on Saturday. The order sequence number for this user is `r instacart$order_number[9]`. This product was the first product added to cart by this user and hasn't been ordered  before. `r instacart$days_since_prior_order[9]` days have past since the last order placed by this user. The evaluation set this order belongs in is `r instacart$eval_set[9]`. 

## Calculate the number of aisles and find out the most frequent

```{r n_aisle}
n_distinct(instacart$aisle) ## number of aisles
instacart %>% 
  count(aisle) %>% 
  top_n(1)
```

There are 134 aisles. Fresh vegetables is the aisle which the most items ordered from.

## Make a plot

The code chunk below makes a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.

```{r aisle_plot, fig.width=15}
instacart %>%
  count(aisle) %>%
  ggplot(aes(x = forcats::fct_reorder(aisle, n), y = n)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8)) +
  labs(x = "Aisle", y = "Number of items", 
       title = "The number of items ordered in each aisle")
```

The number of items ordered in most aisles are below 25,000, although in some aisles, like "fresh fruits" and "fresh vegetables", the number of items ordered are quite large, more than 150,000.

## Make a table 

The code chunk below makes a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r popular_aisle_table}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  top_n(1) %>% 
  rename("the most popular item" = "product_name") %>% 
  select(-n) %>% 
  knitr::kable() 
```

The most popular items aisle “baking ingredients”, “dog food care”, and “packaged vegetables fruits” are Light Brown Sugar, Snack Sticks Chicken & Rice Recipe Dog Treats and Organic Baby Spinach, respectively. 

## Make a table 

The code chunk below makes a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r mean_time}
instacart %>% 
  ## Convert order_dow from numeric to character
  mutate(
    order_dow = recode(order_dow, "0" = "Sunday", "1" = "Monday", "2" = "Tuesday", "3" = "Wednesday", "4" = "Thursday", "5" = "Friday", "6" = "Saturday")
  ) %>%
  mutate(
    order_dow = factor(order_dow, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
  ) %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>%
  group_by(product_name, order_dow) %>% 
  summarize(mean_time = mean(order_hour_of_day, na.rm = TRUE)) %>% 
  spread(key = order_dow, value = mean_time) %>% 
  knitr::kable(digits = 1)
```

On most days of a week, the mean hour at which Pink Lady Apples is ordered is earlier than that and at which Coffee Ice Cream is ordered. 

# Problem 3

## Import the NY NOAA data

```{r ny_noaa_data_import}
data("ny_noaa")
ny_noaa = janitor::clean_names(ny_noaa) ## clean variable names
skimr::skim(ny_noaa)

n_all_missing = ny_noaa %>% 
  filter(is.na(tmax) & is.na(tmin) & is.na(prcp) & is.na(snow) & is.na(snwd)) %>% 
  nrow()
```

Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue.

The NY NOAA data has `r nrow(ny_noaa)` observations and `r ncol(ny_noaa)` variables. Variables for `id`, `tmax` and `tmin` are class `character`. Variables for `prcp`, `snow` and `snwd` are class `integer`. Variable for `date` is class `Date`. The medians of precipitation, snowfall and snow depth during the study period are all 0. The data was collected from 747 unique weather stations. 

There are `r n_all_missing` obseravtions containing no information about temperature, precipitation, snowfall or snow depth, which is about `r round(n_all_missing/nrow(ny_noaa), digits = 2)` of all observaions. The proportion of missing data of `tmax` is `r round(sum(is.na(ny_noaa$tmax))/nrow(ny_noaa), digits = 2)`. The proportion of missing data of `tmin` is `r round(sum(is.na(ny_noaa$tmin))/nrow(ny_noaa), digits = 2)`. The proportion of missing data of `prcp` is `r round(sum(is.na(ny_noaa$prcp))/nrow(ny_noaa), digits = 2)`. The proportion of missing data of `snow` is `r round(sum(is.na(ny_noaa$snow))/nrow(ny_noaa), digits = 2)`. The proportion of missing data of `snwd` is `r round(sum(is.na(ny_noaa$snwd))/nrow(ny_noaa), digits = 2)`. In my opinion, if more than 1/5 of data are missiong, it would become an issue. And whether missing data is an issue also depends on whether the data is randomly missing or non-randomly. 

## Clean the NY NOAA data

```{r ny_noaa_data_clean}
## Convert tmax and tmin to numerical variable
ny_noaa$tmax = as.numeric(ny_noaa$tmax)
ny_noaa$tmin = as.numeric(ny_noaa$tmin)

ny_noaa = ny_noaa %>% 
  ## Create separate variables for year, month, and day
  separate(col = date, into = c("year", "month", "day"), sep = "-") %>% 
  ## Ensure reasonable units
  mutate(
    tmax = tmax/10,
    tmin = tmin/10,
    prcp = prcp/10
  )
```

## The most commonly observed value for snowfall

```{r snowfall}
ny_noaa %>% 
  count(snow) %>% 
  top_n(1)
```

0 is the most commonly observed value for snowfall. Because for most of the time between 1981-01-01 and 2010-12-31, it didn't snow in NY state.

## Make a two-panel plot

The code chunk below makes a two-panel plot showing the average max temperature in January and in July in each station across years.

```{r tmax_jan_july}
ny_noaa %>% 
  filter(month %in% c("01", "07")) %>% 
  group_by(month, year, id) %>%
  summarize(tmax_mean = mean(tmax, na.rm = T)) %>%
  ggplot(aes(x = factor(year), y = tmax_mean)) +
  geom_boxplot() +
  facet_grid(~month) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(x = "Year", y = "The average max temperature", 
       title = "The average max temperature in January and July in each station across years")
```

The average max temperature in January and in July fluctuated a lot across years. In January, except in 1981, 1985, 1989, 1997 and 2006, all other years had outliers. In July, except in 1992, 1997, 1999, 2009 and 2010, all other years had outliers. 

## Make a two-panel plot

The code chunk below makes a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year

```{r tmax_tmin_snowfall}
tmax_tmin = ny_noaa %>% 
  ggplot(aes(x = tmax, y = tmin)) +
  geom_hex() +
  theme(legend.text = element_text(angle = 60, hjust = 1))

snowfall = ny_noaa %>% 
  filter(snow > 0 & snow < 100) %>% 
  group_by(year) %>% 
  ggplot(aes(x = year, y = snow)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

tmax_tmin + snowfall
```

Tmin had the positive relationship with tmax. And The distribution of snowfall values greater than 0 and less than 100 remained almost the same across years. 
